Verifying synchronous distributed algorithms using IVy

### 2018-08-11 paper/paper.tex
nobreak mdframed
proof -> prove

\section{IVy}
IVy is a system designed to verify correctness of infinite-sate systems using the aforementioned deductive approach. It takes as input a specification of an algorithm operating on system-state  and a set of invariants. It proofs the inductiveness of the invariants in respect to the algorithm, or displays a counterexample to induction. If the set of invariants entails the desired properties of the algorithm, and Ivy proved induction, then it was proven that the algorithm is \textit{correct}.

In Ivy, the entity ``calling'' the algorithm in different state configurations is called the ``environment''. The environment conceptually represents the final users of the algorithm. An algorithm or protocol is \textit{correct} or \textit{safe} if it cannot be called by the environment in a state configuration that causes some pre-defined properties (assertions, invariants) not to be satisfied. IVy supports verification of systems that have multiple methods (\textit{actions}) exported to the environment. For example to verify servers that specify connection and disconnection algorithms. This can be thought of as proofing inductiveness of two independent algorithms with respect to the same invariants and system-state.

@@

\subsubsection{Types and declarations}
In IVy one mainly uses \textit{uninterpreted} types. \textit{Uninterpreted} means that the type is an arbitrary set of values with at least one element.
The cardinality of this set is \textit{parameterized}, meaning IVy proofs correctness for an arbitrary amount of values of an \textit{uninterpreted} type.
The number of values of a type that are generated to construct a counter example to induction (CTI) is entirely up to IVy. No instances of these types are manually instantiated.
This is useful for declaring types such as \textit{node}, \textit{process} or \textit{value} in a system.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
type node
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
type node
type value
function value(N:node) -> value
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
function message_received(RECEIVER:node, SENDER:node) -> bool
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
relation message_received(RECEIVER:node, SENDER:node)
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}

@@

To assign values to relations, functions or variables $:=$ is used.
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action our_action(x:node) = {
  failed(x) := true;
}
\end{lstlisting}
\end{mdframed}
Lower case tokens refer to an individual value. In the code above \textit{x} refers to the specific instance of type \textit{node} that has been passed to \textit{our\_action}. The value of the relation \textit{failed(x)} is changed to \textit{true} only for that instance of \textit{node}.

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
failed(X:node) := ~incoming_message(X:node);
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
# Pseudocode only:
foreach x of type node {
  failed(x) := ~incoming_message(x);
}
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
failed(X) := *;
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action random_node_sends_message_to(x:node) = {
  y := *;
  assume ~failed(y);
  send(x, y);
}
\end{lstlisting}
\end{mdframed}

@@

\subsubsection{Invariants}
To specify the properties we want to proof about our algorithm we encode them in special assertions called \textit{invariants}:

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant system.rounds >= 0
invariant system.rounds < 100
\end{lstlisting}
\end{mdframed}
The conjunction of these must hold at all times when an action is not executing, thus they have to be initially true and each of
the \textit{exported} actions must preserve them. However, an invariant is not guaranteed to hold when an action calls another action.

@@

We refer to these condition as ``correctness conditions'', as we say the algorithm is ``correct'' if we can proof it satisfies these conditions.  \textit{Floodset} solves the described problem by defining $msgs$ and $trans$ functions for nodes in the network such that the conditions are satisfied.

@@

\textbf{Formal definition of \textit{Floodset}}:
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $states_i$:
\begin{gather*}
  rounds   \in  \mathbb{N}, \ \mathrm{initially} \ 0 \\
  decision  \in  V \cup \{ unknown \}, \ \mathrm{initially} \ unknown \\
  W \subseteq V, \ \mathrm{initially \ the\  singleton\  set\  consisting\  of\  i's\  initial\  value}
\end{gather*}
\end{mdframed}

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $msgs_i$:
\begin{gather*}
  if\ rounds \le f\ \mathrm{send}\ W\ \mathrm{to\ all\ other\ processes}
\end{gather*}
\end{mdframed}

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $trans_i$:
\begin{gather*}
  rounds\ :=\ rounds+1\\
  \mathrm{let}\ X_j\ \mathrm{be\ the\ message\ from\ j,\ for\ each\ j\ from\ which\ a\ message\ arrives}\\
  W\ :=\ W \cup \bigcup_{j} X_j\\
  \mathrm{if}\ rounds = f+1\ \mathrm{then}\\
  \ \ \   \mathrm{if}\ |W| = 0\ \mathrm{then}\ decision := v,\ \mathrm{where}\ W = {v}\\
  \ \ \   \mathrm{else}\ decision := v_0
\end{gather*}
\end{mdframed}

@@

\section{Specification}
To summarize: we have a parameterized system with $n$ process and $f$ number of faults. We want to show \textit{Floodset} is correct regardless of the parameters.
It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f,\ n < f$''.
To answer this question the algorithm was encoded in the IVy language and then inductive invariants were constructed to proof correctness.

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action round = {
  var iter := node.t.iter.create(0);
  while ~iter.is_end
  {
    call send_messages(iter.val)
    iter := iter.next;
  }
  . . .
}
action send_messages(n:node.t) = {
  node.message(n,V) := node.w(n,V);
  node.incomming_message(O,n) := true;
}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
. . .
line 3: message := *;
line 3: incomming_message := *;
. . .
\end{lstlisting}
\end{mdframed}

IVy non-deterministically assigns to all relations involved in the loop, instead of actually executing the loop.  This is because proving correctness of loops can only be done through inductive loop invariants specified by a human. IVy therefore allows specifying invariants together with a loop.
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
  while ~iter.is_end
  invariant node.message(N,V) = node.w(N,V)
  invariant N <= iter.val -> incomming_message(O,N)
  {
    . . .
  }
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
assert message(N,V) = w(N,V)
assert N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
incomming_message := *
message := *
assume message(N,V) = w(N,V)
assume N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
clean_round_occured := (clean_round_occured |
                        (forall N . ~node.failed_during_round(N)));
if clean_round_occured {
  decision_round := *;
};
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
# lemma 1:
invariant system.clean_round_occured ->
            ((~node.failed(A) & ~node.failed(B)) ->
              (node.w(A, V) = node.w(B, V)))
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
after init {
  def := *;
  default(N) := def;
  w(N, V) := V=node.val(N);
  incomming_message(A, B) := false;
  decision(N, V) := false;
  clean_round_occured := false;
  decision_round := false;
  node.failed_before_round(N) := false;
  node.failed(N) := false;
  node.has_single_value(N) := false;
}
\end{lstlisting}
\end{mdframed}

@@

\noindent A round in the specification is defined as follows:
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action round = {
  require ~node.decision(N,V) & ~decision_round;
  node.failed_before_round(N) := node.failed(N);
  call step1;
  call simulate_failure;
  clean_round_occured := (clean_round_occured |
                          (forall N . ~node.failed_during_round(N)));
  if clean_round_occured {
    decision_round := *;
  };
  call step2;
}
export round;
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action step1 = {
  node.message(A,V) := node.w(A,V) & ~node.failed(A);
  node.incomming_message(O,A) := ~node.failed(A);
}
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action step2 = {
  node.w(N,V) := (node.w(N,V) |
                 (~node.failed(N) &
                  exists J . (node.incomming_message(N,J) &
                               node.message(J,V))));

@@

To recap, a process may fail before or after performing some instance of Step 1 or Step 2 as well as in the middle of performing Step 1.  The presented code for actions \textit{step1} and \textit{step2} already correctly considers the behaviour of nodes that failed before the actions. E.g. \textit{step1} only allows nodes that have not failed to send messages.
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
. . .
node.incomming_message(O,A) := ~node.failed(A);
. . .
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
node.failing_now(N) := *;
assume node.failing_now(N) -> ~node.failed(N);
node.failed_during_round(N) :=
  (node.failed_now(N) | node.failed_during_round(N))
node.failed(N) := (node.failed(N) | node.failed_during_round(N));
assume exists N . ~node.failed(N);
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
node.incomming_message_failing(N,M) := *;
node.incomming_message(N,M) :=
  (node.failed_during_round(M) &
    node.incomming_message_failing(N,M)) |
  (~node.failed_during_round(M) &
    node.incomming_message(N,M));
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
  action simulate_failure = {
    node.failed_during_round(N) := *;
    assume node.failed_during_round(N) -> ~node.failed(N);
    node.failed(N) := (node.failed(N) | node.failed_during_round(N));
    assume exists N . ~node.failed(N);

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant (~node.failed(A) & ~node.failed(B)) ->
            (node.decision(A,V) = node.decision(B,V))
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant node.decision(N,V) -> (~exists O . node.decision(N,O) & O~=V)
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant (exists N,M . node.val(N) ~= node.val(M)) |
          (system.decision_round ->
            (~node.failed(N2) -> node.decision(N2,node.val(N2))))
\end{lstlisting}
\end{mdframed}

>>>>

\section{IVy}
IVy is a system designed to verify correctness of infinite-sate systems using the aforementioned deductive approach. It takes as input a specification of an algorithm operating on system-state  and a set of invariants. It proves the inductiveness of the invariants in respect to the algorithm, or displays a counterexample to induction. If the set of invariants entails the desired properties of the algorithm, and Ivy proved induction, then it was proven that the algorithm is \textit{correct}.

In Ivy, the entity ``calling'' the algorithm in different state configurations is called the ``environment''. The environment conceptually represents the final users of the algorithm. An algorithm or protocol is \textit{correct} or \textit{safe} if it cannot be called by the environment in a state configuration that causes some pre-defined properties (assertions, invariants) not to be satisfied. IVy supports verification of systems that have multiple methods (\textit{actions}) exported to the environment. For example to verify servers that specify connection and disconnection algorithms. This can be thought of as proving inductiveness of two independent algorithms with respect to the same invariants and system-state.

@@

\subsubsection{Types and declarations}
In IVy one mainly uses \textit{uninterpreted} types. \textit{Uninterpreted} means that the type is an arbitrary set of values with at least one element.
The cardinality of this set is \textit{parameterized}, meaning IVy proves correctness for an arbitrary amount of values of an \textit{uninterpreted} type.
The number of values of a type that are generated to construct a counter example to induction (CTI) is entirely up to IVy. No instances of these types are manually instantiated.
This is useful for declaring types such as \textit{node}, \textit{process} or \textit{value} in a system.

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
type node
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
type node
type value
function value(N:node) -> value
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
function message_received(RECEIVER:node, SENDER:node) -> bool
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
relation message_received(RECEIVER:node, SENDER:node)
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}

@@

To assign values to relations, functions or variables $:=$ is used.
\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action our_action(x:node) = {
  failed(x) := true;
}
\end{lstlisting}
\end{mdframed}
Lower case tokens refer to an individual value. In the code above \textit{x} refers to the specific instance of type \textit{node} that has been passed to \textit{our\_action}. The value of the relation \textit{failed(x)} is changed to \textit{true} only for that instance of \textit{node}.

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
failed(X:node) := ~incoming_message(X:node);
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
# Pseudocode only:
foreach x of type node {
  failed(x) := ~incoming_message(x);
}
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
failed(X) := *;
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action random_node_sends_message_to(x:node) = {
  y := *;
  assume ~failed(y);
  send(x, y);
}
\end{lstlisting}
\end{mdframed}

@@

\subsubsection{Invariants}
To specify the properties we want to prove about our algorithm we encode them in special assertions called \textit{invariants}:

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant system.rounds >= 0
invariant system.rounds < 100
\end{lstlisting}
\end{mdframed}
The conjunction of these must hold at all times when an action is not executing, thus they have to be initially true and each of
the \textit{exported} actions must preserve them. However, an invariant is not guaranteed to hold when an action calls another action.

@@

We refer to these condition as ``correctness conditions'', as we say the algorithm is ``correct'' if we can prove it satisfies these conditions.  \textit{Floodset} solves the described problem by defining $msgs$ and $trans$ functions for nodes in the network such that the conditions are satisfied.

@@

\textbf{Formal definition of \textit{Floodset}}:
\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $states_i$:
\begin{gather*}
  rounds   \in  \mathbb{N}, \ \mathrm{initially} \ 0 \\
  decision  \in  V \cup \{ unknown \}, \ \mathrm{initially} \ unknown \\
  W \subseteq V, \ \mathrm{initially \ the\  singleton\  set\  consisting\  of\  i's\  initial\  value}
\end{gather*}
\end{mdframed}

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $msgs_i$:
\begin{gather*}
  if\ rounds \le f\ \mathrm{send}\ W\ \mathrm{to\ all\ other\ processes}
\end{gather*}
\end{mdframed}

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $trans_i$:
\begin{gather*}
  rounds\ :=\ rounds+1\\
  \mathrm{let}\ X_j\ \mathrm{be\ the\ message\ from\ j,\ for\ each\ j\ from\ which\ a\ message\ arrives}\\
  W\ :=\ W \cup \bigcup_{j} X_j\\
  \mathrm{if}\ rounds = f+1\ \mathrm{then}\\
  \ \ \   \mathrm{if}\ |W| = 0\ \mathrm{then}\ decision := v,\ \mathrm{where}\ W = {v}\\
  \ \ \   \mathrm{else}\ decision := v_0
\end{gather*}
\end{mdframed}

@@

\section{Specification}
To summarize: we have a parameterized system with $n$ process and $f$ number of faults. We want to show \textit{Floodset} is correct regardless of the parameters.
It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f,\ n < f$''.
To answer this question the algorithm was encoded in the IVy language and then inductive invariants were constructed to prove correctness.

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action round = {
  var iter := node.t.iter.create(0);
  while ~iter.is_end
  {
    call send_messages(iter.val)
    iter := iter.next;
  }
  . . .
}
action send_messages(n:node.t) = {
  node.message(n,V) := node.w(n,V);
  node.incomming_message(O,n) := true;
}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
. . .
line 3: message := *;
line 3: incomming_message := *;
. . .
\end{lstlisting}
\end{mdframed}

IVy non-deterministically assigns to all relations involved in the loop, instead of actually executing the loop.  This is because proving correctness of loops can only be done through inductive loop invariants specified by a human. IVy therefore allows specifying invariants together with a loop.
\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
  while ~iter.is_end
  invariant node.message(N,V) = node.w(N,V)
  invariant N <= iter.val -> incomming_message(O,N)
  {
    . . .
  }
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
assert message(N,V) = w(N,V)
assert N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
incomming_message := *
message := *
assume message(N,V) = w(N,V)
assume N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
clean_round_occured := (clean_round_occured |
                        (forall N . ~node.failed_during_round(N)));
if clean_round_occured {
  decision_round := *;
};
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
# lemma 1:
invariant system.clean_round_occured ->
            ((~node.failed(A) & ~node.failed(B)) ->
              (node.w(A, V) = node.w(B, V)))
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
after init {
  def := *;
  default(N) := def;
  w(N, V) := V=node.val(N);
  incomming_message(A, B) := false;
  decision(N, V) := false;
  clean_round_occured := false;
  decision_round := false;
  node.failed_before_round(N) := false;
  node.failed(N) := false;
  node.has_single_value(N) := false;
}
\end{lstlisting}
\end{mdframed}

@@

\noindent A round in the specification is defined as follows:
\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action round = {
  require ~node.decision(N,V) & ~decision_round;
  node.failed_before_round(N) := node.failed(N);
  call step1;
  call simulate_failure;
  clean_round_occured := (clean_round_occured |
                          (forall N . ~node.failed_during_round(N)));
  if clean_round_occured {
    decision_round := *;
  };
  call step2;
}
export round;
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action step1 = {
  node.message(A,V) := node.w(A,V) & ~node.failed(A);
  node.incomming_message(O,A) := ~node.failed(A);
}
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action step2 = {
  node.w(N,V) := (node.w(N,V) |
                 (~node.failed(N) &
                  exists J . (node.incomming_message(N,J) &
                               node.message(J,V))));

@@

To recap, a process may fail before or after performing some instance of Step 1 or Step 2 as well as in the middle of performing Step 1.  The presented code for actions \textit{step1} and \textit{step2} already correctly considers the behaviour of nodes that failed before the actions. E.g. \textit{step1} only allows nodes that have not failed to send messages.
\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
. . .
node.incomming_message(O,A) := ~node.failed(A);
. . .
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
node.failing_now(N) := *;
assume node.failing_now(N) -> ~node.failed(N);
node.failed_during_round(N) :=
  (node.failed_now(N) | node.failed_during_round(N))
node.failed(N) := (node.failed(N) | node.failed_during_round(N));
assume exists N . ~node.failed(N);
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
node.incomming_message_failing(N,M) := *;
node.incomming_message(N,M) :=
  (node.failed_during_round(M) &
    node.incomming_message_failing(N,M)) |
  (~node.failed_during_round(M) &
    node.incomming_message(N,M));
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
  action simulate_failure = {
    node.failed_during_round(N) := *;
    assume node.failed_during_round(N) -> ~node.failed(N);
    node.failed(N) := (node.failed(N) | node.failed_during_round(N));
    assume exists N . ~node.failed(N);

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant (~node.failed(A) & ~node.failed(B)) ->
            (node.decision(A,V) = node.decision(B,V))
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant node.decision(N,V) -> (~exists O . node.decision(N,O) & O~=V)
\end{lstlisting}
\end{mdframed}

@@

\begin{mdframed}[nobreak=true, backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant (exists N,M . node.val(N) ~= node.val(M)) |
          (system.decision_round ->
            (~node.failed(N2) -> node.decision(N2,node.val(N2))))
\end{lstlisting}
\end{mdframed}
